---
title: 'Tidy Tuesday: Monster Movies'
author: "Savannah Damiano"
date: "2024-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```

## Homework Instructions

- Have a dedicated Tidy Tuesday Folder
- Have proper script and file structure
- Submit a knitted RMarkdown file
- Save your visual in the output folder
- You must force yourself to learn at least one _NEW_ thing each week (data wrangling, new geom, new table, new way to edit your theme, new package, rmarkdown, etc.)
- You do not have to use the entire weekly dataset 



## Load Libraries
```{r}
library(here)
library(tidyverse)
library(tidytuesdayR)
library(kableExtra)           # for enchancing the appearance of tables
library(magick)               # create, edit, and compose images
library(webshot2)             # takes screenshots of webpages
```

## Load Data

Upload Monster Movies data from github [Monster Movies](https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-10-29/readme.md) !
```{r}
# Option 1: tidytuesdayR package 
## install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2024-10-29')
## OR
tuesdata <- tidytuesdayR::tt_load(2024, week = 44)

monster_movie_genres <- tuesdata$monster_movie_genres
monster_movies <- tuesdata$monster_movies

# Option 2: Read directly from GitHub

monster_movie_genres <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-29/monster_movie_genres.csv')
monster_movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-29/monster_movies.csv')


#view data
glimpse (monster_movie_genres)
glimpse(monster_movies)

#I will only need the monster_movies .csv
```

## Data Wrangle

Goals:

- To make a table

- The table will show horror monster movies with an average rating of 6 or higher, the year they were made, and their genre.

- Also, the table will include their movie poster.

```{r}

monster <- monster_movies %>%
  filter(genres == "Horror",        #filter for only horror movies
         average_rating >= 6)%>%    #filter movies with an average rating of 6 or higher
  arrange(desc(average_rating)) %>% #arrange from highest to lowest rating 
  select(primary_title, year, genres, average_rating)

monster$posterlink <- c(    #add movie posters to dataset
 "https://image.tmdb.org/t/p/original/hFay7CrI4QyL2ZP4Ea9Fg0NpEuO.jpg", 
 "https://image.tmdb.org/t/p/original/9Z4IQsYMXnkY5UT7Tclu2qUYZNM.jpg",
 "https://image.tmdb.org/t/p/original/vNdjSQpNFgI7LLGmERQVLIfGH6G.jpg",
 "https://image.tmdb.org/t/p/original/pAelOfg6nEuOJYcnbPs02rZsZC4.jpg",
 "https://icep.ug/upload/source/Water%20Monster%202.jpg",
 "https://cdn.traileraddict.com/content/unknown/closet-monster.jpg",
 "https://m.media-amazon.com/images/M/MV5BMTU1ODQ3NjIzM15BMl5BanBnXkFtZTcwMTY1MDAwMQ@@._V1_FMjpg_UX1000_.jpg",
 "https://res.cloudinary.com/calartsposters/image/fetch/w_400,f_auto/https://posters.calarts.edu/wp-content/uploads/2022/11/2022_JoshLin_PaulinaGruzdev_MonsterMash.jpg",
 "https://image.tmdb.org/t/p/original/v7WYLLFme3RPyS2i2nz9oWzrTXg.jpg",
 "https://image.tmdb.org/t/p/original/kOcYwbjnEWy7YJJQhmVCb5gyDl2.jpg"
)
```

## Make a Table

[Customize colors in a kable table](https://r-graph-gallery.com/362-customize-colors-in-kableExtra.html#:~:text=In%20order%20to%20have%20a%20gradient%20of%20colors%2C,be%20proportional%20to%20the%20values%20in%20this%20column.)

[HTML Table with kable & kableExtra](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html)

[Designing Tables Video](https://youtu.be/JqUViTDoSEo)
```{r HomeworkTable, message=FALSE, warning=FALSE}

# Function to create HTML image tags from the links I put in my dataframe
create_image_tag <- function(link) {
  paste0('<img src="', link, '" width="75" height="100">')  # Adjust size as needed
}

# Add a new column "image" to my monster data frame 
# I did this so I can use 'sapply'
# 'sapply' applies the create_image_tag function to each posterlink

monster$image <- sapply(monster$posterlink, create_image_tag)


# Create HTML table

movies_table <- kable(monster[, c("primary_title", "year", "genres", "average_rating", "image")],
                                          # using kable- selecting the columns I want in the table
                      format = "html",    # Specify the output format
                      escape = FALSE,     # Allows the HTML to render properly
                      col.names = c("Film", "Release Year", "Genres", "Rating", "Poster")) %>%
                                          # Rename columns for display on table (like labs in ggplot)
  add_header_above(c("Average Rating of the Top 10 Horror Movies" = 5)) %>%   # Add a title
  kable_styling(full_width = FALSE) %>%   # Adjusts table width
  row_spec(0, bold = TRUE, color = "white", background = "black") %>% # Header row styling 
  kable_styling(bootstrap_options = "bordered") %>% #add lines between each column and row
  column_spec(4, color= "white",
              background = spec_color(monster$average_rating[1:10], begin = .04, end = 0.5,
                                      option = "C"))   #creates a color gradient for the 4th column 


#Display my table  
movies_table


#save table to output

save_kable(movies_table, file = here("TidyTuesday", "Week_01_MonsterMovies", "output", "my_table.html"))
```




# What I Learned

- I new how to create/add a new column into a dataset, but I did not know how to manually create and add my own data into each row. I was able to show this new skill by adding in the links for my movie 

  - monster$posterlink <- c()
  

- I learned how to do a HTML table to allow my table to be more interative.

- To make a color gradient on a table column.


# What I Practiced

- I practiced reordering my data so that the movies would be ordered by highest average rating to lowest average rating. I did this by using the arrange(desc() argument. 

- Creating a function.

- Creating a table.
















