---
title: "Week 7A Homework Tidy Tuesday"
author: "Savannah Damiano"
date: "2024-10-10"
output: 
  html_document:
    toc: TRUE                   
    toc_float: TRUE
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```
# Homework Instructions

- You need to make a map **using RMarkdown** with one of the following datasets from past Tidy Tuesdays. 
- Make sure, as usual, that you scripts and output are saved appropriately.

# My Dataset Selection

- I chose the dataset about cats because I love cats! Also, my _indoor_ cat would be mad at me if I didn't (highlighting _indoor_ because this article is about the conservation concerns around outdoor cats)!
- [Cats on the Move](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-31/readme.md)


# Load Libraries
```{r}
library(tidyverse)
library(here)
library(maps)        #libraries to map
library(mapdata)
library(mapproj)
library(leaflet)     #interactive maps
library(htmlwidgets) #to uplaod my leaflet graph to output

```
# Load Data

- Copied methods from [Tidy Tuesday Github Page](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-31/readme.md) 

```{r }
# Get the Data

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the data sets for the week of interest

# Either ISO-8601 date or year/week works!

tuesdata <- tidytuesdayR::tt_load('2023-01-31')
tuesdata <- tidytuesdayR::tt_load(2023, week = 5)

cats_uk <- tuesdata$cats_uk
cats_uk_reference <- tuesdata$cats_uk_reference

# Or read in the data manually

cats_uk <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk.csv')
cats_uk_reference <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk_reference.csv')

```
# Inspect Data

- I need to look at and inspect the data so I can know what I am getting myself into!

```{r}
view(cats_uk)
view(cats_uk_reference)
head(cats_uk)
head(cats_uk_reference)

#looking at both data sets in a couple of different ways
#also reminding myself that we are making a map
```
# Data Modification 

I want to look at the movement of two kitty cats, Casper and Poppet. 

```{r}

cat_names <- cats_uk %>%               #renamed the data set 
  select(tag_id) %>%                   #selecting the tag_id column
  reframe(name = unique(tag_id)) %>%   
  # I originally thought to use summarize() but R told me to use reframe
  # Since there are a lot of cats, I will make a new column so I can see all the tag_id names so I can easily pick a cat
  arrange(name)
  # We learned this a couple weeks ago! It arranges the names into alphabetical order. 

#Since it's spooky season, it is only law that I pick Casper!
casper <- cats_uk %>%
  filter(tag_id == "Casper-Tag")   #filters to only see Casper-Tag data

#Poppet is fun to say so we will chose that friend for out second map.
poppet <- cats_uk %>%
  filter(tag_id == "Poppet-Tag")   #filters to only see Poppet-Tag data


```

# Creating a Map Using the Leaflet Package 

Map Specifics:

-   I used the leaflet package.

-   I gathered information about the package through the following websites:

    -[Introduction to leaflet](https://rstudio.github.io/leaflet/articles/leaflet.html#:~:text=Create%20a%20map%20widget%20by%20calling%20leaflet%28%29.%20Add,desired.%20Print%20the%20map%20widget%20to%20display%20it.)
    
    -[The Map Widget](https://rstudio.github.io/leaflet/articles/widget.html)


#### Casper Map
```{r HomeworkMap1}

caspermap <- leaflet(casper) %>%
  addTiles() %>%  #adds a tile layer
  addPolylines(lng = ~location_long, lat = ~location_lat, opacity = 0.5, color = "black") %>%
  # adds lines to connect the dots of where the cat travels
  #lng = ~longitude name in df
  #lat = ~latitude name in df
  #opacity changes the transparency 
  #changing the colors of lines
  addCircles(lng = ~location_long, lat = ~location_lat,  opacity = 0.5, color = "orange") %>%
  #adds the circles on the map
  #changing colors of circles
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "feet",
    primaryAreaUnit = "sqfeet")
  #add a interactive measuring tape (in feet) in the bottom left corner

caspermap       #view map

saveWidget(caspermap, file = "HomeworkMap1.html")
#save map

#ummm... Casper is kind of boring.... I guess I should try another cat?
#Wait! If you zoom in a little more on Hayle city, Capser is kind of fun. 
```

#### Poppet Map
```{r HomeworkMap2}
poppetmap <- leaflet(poppet) %>%
  addTiles() %>%  #adds a tile layer
  addPolylines(lng = ~location_long, lat = ~location_lat, opacity = 0.8, color = "hotpink") %>%
  addCircles(lng = ~location_long, lat = ~location_lat,  opacity = 0.5, color = "purple") %>%
  addMeasure(
    position = "bottomleft",
    primaryLengthUnit = "feet",
    primaryAreaUnit = "sqfeet")

# refer to previous r chuck for comments 

poppetmap      #view map

saveWidget(poppetmap, file = "HomeworkMap2.html")
#save map

#Okay Poppet! 
```


# Dataset Credit
- Kays R, Dunn RR, Parsons AW, Mcdonald B, Perkins T, Powers S, Shell L, McDonald JL, Cole H, Kikillus H, Woods L, Tindle H, Roetman P (2020) The small home ranges and large local ecological impacts of pet cats. Animal Conservation. doi:10.1111/acv.12563
- McDonald JL, Cole H (2020) Data from: The small home ranges and large local ecological impacts of pet cats [United Kingdom]. Movebank Data Repository. doi:10.5441/001/1.pf315732














