---
title: "Homework 13a: Interactive Coding"
author: "Savannah Damiano"
date: "2024-12-8"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```


## Homework

- You have a set of 4 .csv files in data/homework. 

- Each of these files is a time series of temperature and light data collected in tide pools in Oregon by Jenn Fields. 

- Your goal is to bring in all 4 files and calculate the mean and standard deviation of both temperature (Temp.C) and light (Intensity.lux) for each tide pool. 

- Use both a for loop and map() functions in your script. (Basically, do it twice). 

- Due Tuesday at 1pm.

## Load Libraries
```{r}
library(tidyverse)
library(here)
```

## Load Data
```{r}
# I am only loading the .csv here so I can view the data before I start the HW.

#TP1<-read_csv(here("Week_13","Data", "homework", "TP1.csv"))

#TP2<-read_csv(here("Week_13","Data", "homework", "TP2.csv"))

#TP3<-read_csv(here("Week_13","Data", "homework", "TP3.csv"))

#TP4<-read_csv(here("Week_13","Data", "homework", "TP2.csv"))
```


## for loop

```{r}
#Step 1: Tell .R the location of the files on the computer
Tpath <- here("Week_13","Data", "homework")         #location of folder
Tfiles <- dir(path = Tpath, pattern = ".csv")       #every file in the folder that has a .csv pattern
Tfiles                                              #check to make sure all 4 .csv are in 'files' value

#Step 2: Pre-allocate space for a loop & calculate mean and sd
loopdata<- tibble( filename= rep(NA, length(Tfiles)),    #column name for PoolID. length is the number of files we imported
                   mean_temp= rep(NA, length(Tfiles)),   #column name for mean temp calc.
                   sd_temp= rep(NA, length(Tfiles)),     #column name for sd temp calc.
                   mean_light= rep(NA, length(Tfiles)),  #column name for mean light calc.
                   sd_light= rep(NA, length(Tfiles)))     #column name for sd light calc.
                   
loopdata #check df

#Step 3: Write a basic code to calculate mean to check that the code works
# We will test by reading in the first file [1] 
# paste0 is a function used to combine strings without any space or separator between them
# "/" used to separate the directory path and the filename
raw_data <- read_csv(paste0(Tpath,"/",Tfiles[1]))   
head(raw_data) #check df

mean_temp <- mean(raw_data$Temp.C, na.rm = TRUE)   #calculate the mean of temp.c
mean_temp

#Step 4: Turn it into a for loop
for (i in 1:length(Tfiles)) {       #loop through the number of files in Tfiles
  raw_data <- read_csv(paste0(Tpath,"/",Tfiles[1]))  #read in .csv in the Tfiles index.  

  loopdata$filename[i] <-Tfiles[i]  #put the file names in this column
  
  loopdata$mean_temp[i] <- mean(raw_data$Temp.C, na.rm = TRUE)   #calculate the mean temp (for each file) and place it in the corresponding row
  loopdata$sd_temp[i] <- sd(raw_data$Temp.C, na.rm = TRUE)       #calculate the sd temp (for each file) and place it in the corresponding row
  loopdata$mean_light[i] <- mean(raw_data$Intensity.lux, na.rm = TRUE) #calculate the mean light (for each file) and place it in the corresponding row
  loopdata$sd_light[i] <- sd(raw_data$Intensity.lux, na.rm = TRUE) #calculate the sd light (for each file) and place it in the corresponding row
}

loopdata #check df
```

## map()

```{r}

#Step 1: Tell .R the location of the files on the computer
TPpath<- here("Week_13","Data", "homework")
files<- dir(path = TPpath, pattern = ".csv", , full.names =TRUE)
files #check to make sure all 4 .csv are in 'files' value

#Step 2: Read in the files using map()
mapdata <- files %>%
  set_names() %>%                      # this set's the .id of each list to the file name 
  map_df(read_csv, .id = "filename")%>% # asks .r to map the data into a df and to label the column with the filenames 'filenames'. And saves the entire path name.

#Step 3: Calculate mean and standard deviation of both temperature (Temp.C) and light (Intensity.lux) for each tide pool
group_by(PoolID) %>%  #group by Tide Pool
summarise(
    mean_temp = mean(Temp.C, na.rm = TRUE),        #calculate mean temp, remove NA
    sd_temp = sd(Temp.C, na.rm = TRUE),            #calculate standard deviation, remove NA
    mean_light = mean(Intensity.lux, na.rm = TRUE),#calculate mean temp, remove NA
    sd_light = sd(Intensity.lux, na.rm = TRUE))    #calculate standard deviation, remove NAA

mapdata #check df

```















