---
title: "Homework 9B"
author: "Savannah Damiano"
date: "2024-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```

# Homework Instructions

- Working with interidal data
- Words: clean up issues with the character data
- Factors: reorder your factors into orders that make sense
- Fix column headers
- Use the Janitor package
- Make a plot: One of the **axes** must be a **factor**
- Make sure graphs go in chronological/ hierarchical/ logical order

# Load Libraries

```{r}
library(here)
library(tidyverse)
library(janitor)              #helps clean data
library(tidytext)             #for text mining and making tidy text
library(PNWColors)            #colors
```

# Load Data

```{r}
# We learned about a common factor booby-trap when reading in data
# We have learned to read in data as 'read_csv()' so the strings will be read in as characters instead of factors
intertidal <- read_csv(here("Week_09", "Data", "intertidaldata.csv"))
view(intertidal)

intertidal_lat <- read_csv(here("Week_09", "Data", "intertidaldata_latitude.csv"))
view(intertidal_lat)
```

# Data Wrangling 

I will use the intertidaldata.csv data for this homework.

```{r}
#Clean the column names and random characters in the quadrat column 

intertidal_clean <- intertidal %>%
  clean_names() %>%  # coverts column names to consistent format 
   mutate(quadrat = str_remove_all(quadrat, "[\\.0-9\\s]")) #remove periods, whitespace, and numbers from quadrat column

view(intertidal_clean)
# view the new data to make sure it is clean 

tidal_plot <- intertidal_clean %>%
  filter(site == "Scripps") %>%               #only want data from scripps 
  pivot_longer(cols = c(bare_rock: anemone),  # % cover data
               names_to = "Cover",
               values_to = "Percent") %>%
  select(c(site, transect, quadrat, Cover, Percent)) %>%  #select out the data I want 
  mutate(quadrat = factor(quadrat)) %>%    #makes quadrat a factor 
  droplevels()                              # drops extra levels



#After plotting I realized that there was two words for low and two words for mid. We need to fix this so we only have one word to represent high, mid, and low. 

unique(tidal_plot$quadrat)   
#check the unique values of quadrat to view the inconsitancy 

#I realized the values had whitespace that needed to be removed so i went back to the 'mutate(quadrat = str_remove_all()' and added a whitespace remover

#I also saw that the legend ordered the quadrats from High- Low- Mid, but it needs to be High-Mid-Low for ease of our reader (& per our homework instructions)
tidal_plot$quadrat <- factor(tidal_plot$quadrat, levels = c("Low", "Mid", "High"))

```

# Plot

```{r Homework9B}
pal <-pnw_palette("Sailboat",6,type="discrete")   #add a color palette for plot

ggplot(tidal_plot,
       aes(x = Cover,
           y = Percent,
           fill = quadrat)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percent Cover by Quadrat at Scripps",
       x = "Cover Type",
       y = "Percent Cover",
       fill = "Tide") +         #change legend title
  scale_fill_manual(values=pal) +  #add a color palette 
  theme_bw()+                   #change theme
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + #turn axis text 
  scale_x_discrete(
    labels = c("algae" = "Algae",
               "anemone" = "Anemone",
               "bare_rock" = "Bare Rock",
               "gooseneck_barnacles" = "Gooseneck Barnacles",
               "large_barnacles" = "Large Barnacles",
               "mussels" = "Mussels",
               "small_barnacles" = "Small Barnacles"))  # New x axis labels

  

```













