---
title: "Homework 9A"
author: "Savannah Damiano"
date: "2024-10-23"
output: 
  html_document:
    toc: TRUE                   
    toc_float: TRUE
    theme : journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```
# Homework Instructions

"Use any dataset you would like and make two useful functions. One that is a plot and one that is not. Turn in your functions using RMarkdown with at least 2 examples per function showing its flexibility."

# Load Libraries
```{r }
library(here)
library(tidyverse)    
library(PNWColors)   #color palette 
```

# Load Data

```{r }
EnviroData<-read_csv(here("Week_05","Data", "site.characteristics.data.csv"))

glimpse(EnviroData)  #make it visible on .rmd
```

# Non-Plot Function 

A function to calculate the mean of a specific parameter across different sites.
```{r}

# Function to calculate mean of a specific parameter

calculate_mean <- function(data, parameter) {
 
   # Filter data for the specified parameter
  filtered_data <- data %>% filter(parameter.measured == parameter)
  
  # Calculate mean by site
  mean_values <- filtered_data %>%
    group_by(site.name = site.letter) %>%   #group by site
    summarise(mean_value = mean(values, na.rm = TRUE)) #find mean and ignore NA
  
  return(mean_values)  #return the resulting means

}

#Test Functions

#Example 1:
#Calculate mean for light
mean_light <- calculate_mean(EnviroData, "light")
glimpse(mean_light) #make it visible on .rmd

#Example 2
#Calculate mean for temperature
mean_temp <- calculate_mean(EnviroData, "temp")
glimpse(mean_temp) #make it visible on .rmd
```

# Plot Function 

A function to plot the EnviroData data set based on a specified parameter.

```{r 1HomeworkPlot9a}

# Example function to plot data
plot_parameter <- function(data = EnviroData, parameter) {
  
  # Filter the data for the specified parameter
  filtered_data <- data %>% 
    filter(parameter.measured == parameter)
  
  # Create the plot
  ggplot(filtered_data, aes(x = name, y = values, fill = parameter.measured)) +
    geom_bar(stat = "identity", position = "dodge") +
                #bar plot
                #stat = "identity" : use actual values provided in data
                #position = "dodge" : so the bars aren't stacked on top of each other
    theme_minimal() +
                #background theme
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
                #adjust x axis titles so they do not overlap
     scale_fill_manual(values = pnw_palette("Starfish", 6))
                #change color of bars
}

#Test
#Example 1
#light across 6 sites
plot_parameter(EnviroData, "light")
```


```{r 2HomeworkPlot9a}

#Test
#Example 2
#algal cover across 6 sites
plot_parameter(EnviroData, "algal.cover")
```





