---
title: "Tidy Tuesday Week 4"
subtitle: "Bob's Burgers Episode"
author: "Savannah Damiano"
date: "2024/11/22"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE,
                       fig.path = "../Output/")
options(htmltools.dir.version = FALSE)
```
## Homework Instructions

- Have a dedicated Tidy Tuesday Folder
- Have proper script and file structure
- Submit a knitted RMarkdown file
- Save your visual in the output folder
- You must force yourself to learn at least one _NEW_ thing each week (data wrangling, new geom, new table, new way to edit your theme, new package, rmarkdown, etc.)
- You do not have to use the entire weekly dataset 
---

## Load Libraries
```{r}
library(here)            #to simplify file paths
library(tidyverse)       #data manipulation, visualization and analysis
library(tidytuesdayR)    #for weekly datasets
library(xaringan)        #create presentations 
library(ggridges)        #create ridge plots

```


---

## Load Data

Upload data from [Bob's Burgers Episodes](https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-11-19/readme.md) !

```{r}
tuesdata <- tidytuesdayR::tt_load(2024, week = 47)

episode_metrics <- tuesdata$episode_metrics

head(episode_metrics)

```


---
## Our {little tiny} Data Wrangling

```{r}
episode_metrics<- episode_metrics %>%
  filter(episode != 23)

#No matter what I did, I could not get episode 23 data to show on my plot
#So we cut it :)

```

---
--- 


## Plotting the Data  
```{r, fig.show='hide'}
plot <- ggplot(episode_metrics, aes(x = avg_length, 
                            y = factor(episode), 
                            fill = factor(episode))) +
  #I set episode as a factor so I could control the level order later on
  geom_density_ridges(scale = 2, alpha = 0.8, size = 0.7) +  
  # Adjust the scale for density
  theme_ridges() +  
  # Use the theme from ggridges
  labs(title = "Average Length by Episode",
       x = "Average Length (minutes)",
       y = "Episode Number") +
  theme(legend.position = "none", # hide legend
        axis.title.x = element_text(hjust = 0.5),  # Center x-axis label
        axis.title.y = element_text(hjust = 0.5)) +# Center y-axis label
  scale_y_discrete(limits = rev(levels(factor(episode_metrics$episode)))) 
   # Reverse y-axis order
 
```
---
## Bob's Burgers Episode Plot 
```{r Homework, echo=FALSE, fig.height=7, fig.width=9}
plot

ggsave(here("TidyTuesday","Week_04_BobsBurgers","Output","Homework.png"))
```

---

## I learned!

- How to make xaringan slides!
  - I gathered information from [Bookdown](https://bookdown.org/yihui/rmarkdown/xaringan.html), [Yihui's Xaringan Github Page](https://github.com/yihui/xaringan?tab=readme-ov-file), and by looking at Dr. Silbiger's weekly presentation's code. 

- I made my first ridgeline plot (I think this is my first).So I got to learn a new geom. 
  - I gathered information from [The R Graph Gallery](https://r-graph-gallery.com/ridgeline-plot.html). 
--
