---
title: 'Tidy Tuesday : Bob''s Burgers'
author: "Savannah Damiano"
date: "2024-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```

## Homework Instructions

- Have a dedicated Tidy Tuesday Folder
- Have proper script and file structure
- Submit a knitted RMarkdown file
- Save your visual in the output folder
- You must force yourself to learn at least one _NEW_ thing each week (data wrangling, new geom, new table, new way to edit your theme, new package, rmarkdown, etc.)
- You do not have to use the entire weekly dataset 


## Load Libraries
```{r}
library(here)
library(tidyverse)
library(tidytuesdayR)
library(xaringan)
library(ggridges)
```

## Load Data

Upload data from [Bob's Burgers Episodes](https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-11-19/readme.md) !

```{r}
episode_metrics <- tuesdata$episode_metrics

head(episode_metrics)
```

## Plot

```{r}
ggplot(episode_metrics, aes(x = avg_length, y = factor(episode), fill = factor(episode))) +
  #I set episode as a factor so I could control the level order later on
  geom_density_ridges(scale = 2) +  # Adjust the scale for density
  theme_ridges() +  # Use the theme from ggridges
  labs(title = "Average Length by Episode",
       x = "Average Length",
       y = "Episode") +
  theme(legend.position = "none")+  # hide legend
  scale_y_discrete(limits = rev(levels(factor(episode_metrics$episode))))   # Reverse y-axis order
```





















