---
title: "Tidy Tuesday: Democrcy and Dictatorship"
author: "Savannah Damiano"
date: "2024-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```
## Homework Instructions

- Have a dedicated Tidy Tuesday Folder
- Have proper script and file structure
- Submit a knitted RMarkdown file
- Save your visual in the output folder
- You must force yourself to learn at least one _NEW_ thing each week (data wrangling, new geom, new table, new way to edit your theme, new package, rmarkdown, etc.)
- You do not have to use the entire weekly dataset 


## Load Libraries
```{r}
library(here)
library(tidyverse)
library(tidytuesdayR)
library(janitor)
library(gganimate)       # create animated graphs
library(transformr)      # help with transformation with animations
library(gifski)          # help create gifs
```

## Load Data
```{r message=FALSE, warning=FALSE}
# Option 1: tidytuesdayR package 
## install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2024-11-05')
## OR
tuesdata <- tidytuesdayR::tt_load(2024, week = 45)

democracy_data <- tuesdata$democracy_data

# Option 2: Read directly from GitHub

democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-05/democracy_data.csv')

glimpse(democracy_data)

```

# Data Wrangle 


```{r}
clean_data <- democracy_data %>%
 clean_names() %>%                 #janitor package to clean names
 filter(!is.na(monarch_name)) %>%  #filters out rows in this column that have na 
 filter(is_monarchy == TRUE) %>%
 select(country_name, year, regime_category, is_monarchy, monarch_name) %>%     #select out the columns I want
 drop_na(country_name, year, regime_category, is_monarchy, monarch_name)         #drop na
 
#Assign a unique ID to each monarch in each country
plot_data<-clean_data %>%
  group_by(country_name, year)%>%
  mutate(monarch_ID = as.factor(monarch_name))%>%
  filter(country_name %in% c("Belgium", "Bhutan", "Jordan", "Nepal", "Tonga", "Turks and Caicos")) %>%
  ungroup()

# Summarize the data to get the count of each regime
top_regimes <- plot_data %>%
  group_by(regime_category) %>%       # Group by regime category
  summarise(count = n()) %>%          # Count the occurrences of each regime
  arrange(desc(count)) %>%            # Arrange by count in descending order
  top_n(4, count)                     # Select the top 4 regimes

# Filter the dataset to include only the top 4 regimes
filtered_data <- plot_data %>%
  filter(regime_category %in% top_regimes$regime_category)

```

## Plot

```{r Homework}

animated_plot <- ggplot(filtered_data, aes(x = factor(year), fill = regime_category)) +
  geom_bar(stat = "count", position = "dodge", width = 5.0) +
   facet_wrap(~ country_name, scales = "free_y") +  # Facet by country
  labs(title = "Regimes Over Time for ", 
       subtitle = "Year: {round(frame_time)}", #rounds year to whole number
       x = "Year", 
       y = "Count of Regimes") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = seq(min(filtered_data$year), max(filtered_data$year), by = 20)) +  # Show every 2nd year
  transition_time(year) +           # Animate by year
  ease_aes('linear')                # Make the animation smooth


anim_save("animation.gif", animation = animated_plot)





```

# What I Learned 

I learned how to animate a graph! I know she is not cute. I struggled to figure out how to best visualize this data with animation (I was super determined to teach myself animation). 

















