---
title: 'Tidy Tuesday: ISO Country Codes'
author: "Savannah Damiano"
date: "2024-11-18"
output: 
  ioslides_presentation:
    widescreen: true   # Use widescreen aspect ratio
  
    
#I planned to have the individual slides scroll so you could see all the text, however when I did this, the plot would not appear on the slide anymore. So I desided to ditch the idea for the sake of my sanity. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```

## Homework Instructions

- Have a dedicated Tidy Tuesday Folder
- Have proper script and file structure
- Submit a knitted RMarkdown file
- Save your visual in the output folder
- You must force yourself to learn at least one _NEW_ thing each week (data wrangling, new geom, new table, new way to edit your theme, new package, rmarkdown, etc.)
- You do not have to use the entire weekly dataset 



## Load Libraries
```{r}
library(tidytuesdayR)
library(tidyverse)
library(here)
library(maps)        #world dataset
library(mapdata)
```

## Load Data
```{r}
# Option 1: tidytuesdayR package 
## install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2024-11-12')
## OR
tuesdata <- tidytuesdayR::tt_load(2024, week = 46)

countries <- tuesdata$countries
country_subdivisions <- tuesdata$country_subdivisions
former_countries <- tuesdata$former_countries

# Option 2: Read directly from GitHub

countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/countries.csv')
country_subdivisions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/country_subdivisions.csv')
former_countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/former_countries.csv')


# look at data using environment 
```

## Data Wrangling 

```{r}
iso_data <- countries %>%
  select(numeric,name)         
#select out the numeric and the country names
world_map <- map_data("world")   #get world map data

#There are a lot of country name differences between the df
#Need to find all the differences so I can fix them
setdiff(iso_data$name, world_map$region)
setdiff(world_map$region, iso_data$name)

```

```{r}

#fix country names to match both data sets 
#so all countries show up on plot 

world_data <- map_data("world") %>%
  mutate(region = case_when(
    region == "USA" ~ "United States",
    region == "Turkey" ~ "Türkiye",
    region == "UK" ~ "United Kingdom",
    region == "Syria" ~ "Syrian Arab Republic",
    region == "Democratic Republic of the Congo" ~ "Congo, The Democratic Republic of the",
    region == "Iran" ~ "Iran, Islamic Republic of",
    region == "Republic of Congo" ~ "Congo",
    region == "Czech Republic" ~ "Czechia",
    region == "Taiwan" ~ "Taiwan, Province of China",
    region == "North Korea" ~ "Korea, Republic of", 
    region == "Ivory Coast" ~ "Côte d'Ivoire",
    region == "Vietnam" ~ "Viet Nam",
    region == "Russia" ~ "Russian Federation",
    region == "South Korea" ~ "Korea, Democratic People's Republic of",
    region == "Bolivia" ~ "Bolivia, Plurinational State of",
    region == "Venezuela" ~ "Venezuela, Bolivarian Republic of",
    region == "Moldova" ~ "Moldova, Republic of",
    region == "Tanzania" ~ "Tanzania, United Republic of",
    region == "Laos" ~ "Lao People's Democratic Republic",
    TRUE ~ region
    
  ))

```

```{r}
world_map<- world_data %>%
  left_join(iso_data, by= c("region" = "name"))  
#join datasets together based on the country name
#since the datasets have different column names, I specified the names inside the join.
```

## Make a Map

```{r}

plot <- ggplot(world_map) +
  geom_map(aes(map_id = region, fill = numeric), map = world_map) + 
  #creates map by using region column to match countries
  #fill =numeric to   color the countries by their iso code
  
  expand_limits(x = world_map$long, y = world_map$lat) +
  # makes sure the map covers all counties
  
  scale_fill_gradientn(colors = c("lightblue","darkblue")) +
  #defines the color gradient for the iso codes
  
  theme_void() +                       #theme to make the map look cleaner
  labs(title = "ISO Numeric Codes per Country", #labels
       fill = "ISO Numeric Code") +
  theme(legend.position = "bottom")    #places legend below plot
```

## Map
```{r}
print(plot)
```


## What I Learned

- How to make my RMarkdown into a slide show: [isoslides](https://bookdown.org/yihui/rmarkdown/ioslides-presentation.html#custom-templates-2), and how to make it [scroll](https://stackoverflow.com/questions/53064555/how-to-make-scrollable-slides-in-an-ioslides-presentation-with-rmarkdown). However, I could not get the scroll to work.

- How to join two data sets without the column names matching. During my data wrangling, I used the code -  left_join(iso_data, by= c("region" = "name")) . "region" is from the world_map dataframe and "name" is from the iso_data dataframe. 

- How to search for mismatched words between two drataframes.

## What I Practiced

- I realized I have not practiced making a map outside of the one we did in lecture. I figured this dataset is a perfect opportunity to practice this skill! 

- Renaming a million words to match across two dataframes :)
