---
title: "Tidy Tuesday Week 5"
subtitle: "U.S. Customs and Border Protection (CBP) Encounter Data"
author: "Savannah Damiano"
date: "2024-11-30"
output: 
  html_document:
    toc: TRUE                   
    toc_float: TRUE
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning=FALSE,
                      fig.path = "../Output/")
```

## Homework Instructions

- Have a dedicated Tidy Tuesday Folder
- Have proper script and file structure
- Submit a knitted RMarkdown file
- Save your visual in the output folder
- You must force yourself to learn at least one _NEW_ thing each week (data wrangling, new geom, new table, new way to edit your theme, new package, rmarkdown, etc.)
- You do not have to use the entire weekly dataset 


## Load Libraries
```{r}
library(here)            #to simplify file paths
library(tidyverse)       #data manipulation, visualization and analysis
library(tidytuesdayR)    #for weekly datasets
library(janitor)         #to help clean data
library(PNWColors)       #to add colors

```

## Load Data

Upload data from [U.S. Customs and Border Protection (CBP) Encounter Data](https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-11-26/readme.md) from the Tidy Tuesday GitHub page!

```{r}
tuesdata <- tidytuesdayR::tt_load(2024, week = 48)

cbp_resp <- tuesdata$cbp_resp
cbp_state <- tuesdata$cbp_state

head(cbp_resp)
head(cbp_state)
```

## Data Wrangling 

```{r}
tidydata <- cbp_state %>%
  clean_names() %>%         #janitor package cleaning data names
  group_by(month_abbv, state) %>%                #select out these columns
  summarize(total_encounters = sum(encounter_count)) %>% # Sum encounters for each state
  arrange(desc(total_encounters)) %>%  # Sort by total encounters in descending order
  slice_head(n = 5) # Select the top 5 states

# I noticed the months on the x axis were out of order
# 'month_abbv' needs to be an ordered factor with correct month order
month_order <- c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT", "NOV", "DEC")
tidydata$month_abbv <- factor(tidydata$month_abbv, levels = month_order, ordered = TRUE)


```


## New Data Wrangling Technique I Learned 
```{r}
spread_data <- tidydata %>%
  spread(key = month_abbv, value = total_encounters, fill = 0)

#spread is a function that can be used to spread out or reshape the data. 

#key = month_abbv - specifies the column containing the months will be used as new columns in the new df.

#value = total_encounters - tells R that this column will be placed in the new columns by the month variable. Each state in the df will have it's our corresponding total encounter number. 

#fill = 0 - the fill argument is used to fill in missing values. If for a particular state and month combination there is no data (i.e., the value is NA), it will be filled with 0. 

#I am a very visual person, so this technique helped visualize the data for each state and month easier. 

spread_data
```


## Plot 
```{r, HomeworkPlot}
ggplot(tidydata, aes(x = month_abbv, y = total_encounters, fill = state)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bar chart
  theme_minimal() +     #add theme
  labs(                 #change titles
    title = "Total Encounters at the U.S. Border by State and Month",
    subtitle = "Data from 2020 to 2024",
    x = "Month",
    y = "Total Encounters",
    fill = "State Legend",              
    caption = "Source: Tidy Tuesday Dataset (U.S. Customs and Border Protection Encounter Data)"
  ) +
  scale_fill_manual(
    values = c("AZ" = "#5d74a5", "CA" = "#b0cbe7", "FL" = "#fff0b4", "NM" = "#eba07e", "TX" = "#a45851"),  # Custom colors
    labels = c("Arizona", "California", "Florida", "New Mexico", "Texas")  # Custom x axis legend labels
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels for readability
   plot.caption = element_text(hjust = 0.5))  # Center the caption
```

## What I Learned!

- I learned a new geom [Stacked Barplot](https://www.statology.org/stacked-barplot-in-r/)!

- I learned a new argument in the tidyr package [spread()](https://www.statology.org/spread-function-in-r/)!

















